Personal Activity Monitorring Report, single device from October-November 2012
========================================================

This Knitted report is for the Reproducible Research class Homework Assignment #1. This report is an analysis of Peronal Activity
of the number of steps collected over five minute intervals from a single device/person over a two month period from October through
November of 2012.

Loading and preprocessing the data
========================================================

Show any code that is needed to: 
*(1)*  Load the data (i.e.  read.csv() ), and
*(2)*  Process/transform the data (if necessary) into a format suitable for your analysis.

```{r echo = TRUE}

#opts_chunk$set(echo = TRUE, results = "show", cache = TRUE)

#Download the Activity data if it doesn't already exist
#setwd("C://Users//akeller.HARDEN//Documents//GitHub//RepData_PeerAssessment2")
internetSource <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
fullzipfilename <- file.path(getwd(), "activity.zip")
fullfilename <- file.path(getwd(), "activity.csv")
if (!file.exists(fullzipfilename)){
  #print("file not found locally, so dowload from website")
  download.file(internetSource 
                , destfile=fullzipfilename)
}else{
  #print("data previously downloaded and exists locally")
}

#unzip contents to get single .csv file, 
# but only if unzipped .csv file doesn't already exist
if (!file.exists(fullfilename)){
  unzip(fullzipfilename)
}

#load data into R
if(!exists("activity")){
  activity <- read.csv(fullfilename, header = TRUE)
} else {print("activity already loaded")}

activityorig <- activity  #copy off original to later compare to newly copied object

#pad new intervalmod variable with leading zeros so that it is consistantly HHMM (military hrs)
factor(activity$intervalmod)
activity$intervalmod <- sprintf("%04d",activity$interval)

# add a datimetime variable as a POSIXt class which combines date and hour/minutes from interval
activity$datetime <- strptime(
  paste(
    activity$date
    ,activity$intervalmod) 
    ,format="%Y-%m-%d %H%M")

#sum values by Day and by Interval
byday <- tapply(activity$steps,activity$date,sum, na.rm = TRUE )
byday <- aggregate(activity$steps,list(activity$date), sum,na.rm = TRUE )
colnames(byday)[1]<-"day"
colnames(byday)[2]<-"steps"
# summary(byday)

davg <- mean(byday$steps, na.rm=TRUE)
dmed <- median(byday$steps, na.rm=TRUE)



byinterval <- tapply(activity$steps,activity$intervalmod,sum, na.rm = TRUE )
byinterval <- aggregate(activity$steps,list(activity$intervalmod), sum,na.rm = TRUE )
#merge byinterval with stepsmod to replace NULL values in stepsmod
#replaced with average steps for that timeinterval period rounded to the nearest whole numbers
#get interval means and put in a dataframe
colnames(byinterval)[1]<-"interval"
colnames(byinterval)[2]<-"steps"

#need to count number of missings. Used in output later
nMissingCount <- sum(is.na(activity$steps)
    |is.na(activity$date)
    |is.na(activity$interval)
    )

#put interval mean in stepsmod where values are NA in steps. otherwise use actual steps 
activity$stepsmod <-  ifelse(
  is.na(activity$steps)
  , round(byinterval$steps[match(activity$intervalmod, byinterval$intervalmod)], digits=0)
  , activity$steps
  )

activitynew <- activity[,c(6,2,3)]
colnames(activitynew)[1] <- "steps" #rename stepsmod back to steps

bydaynew <- tapply(activitynew$steps,activitynew$date,sum, na.rm = TRUE )
bydaynew <- aggregate(activitynew$steps,list(activitynew$date), sum,na.rm = TRUE )
colnames(bydaynew)[1]<-"day"
colnames(bydaynew)[2]<-"steps"

#byday <- tapply(activity$steps, activity$date, sum)
davgnew <- mean(bydaynew$steps, na.rm=TRUE)
dmednew <- median(bydaynew$steps, na.rm=TRUE)


#AverageDailySteps <- davg$steps
# activity$daytype <- ""
# factor(activity$daytype)
activity$daytype <- factor(ifelse (
  weekdays(activity[,"datetime"])=="Saturday"   #if Saturday or Sunday
  | weekdays(activity[,"datetime"])=="Sunday"
  , "Weekend"               #then Weekend
  ,"Weekday"))

```


A. What is mean total number of steps taken per day?
========================================================

For this part of the assignment, you can ignore the missing values in the dataset.
*(A1)*  Make a histogram of the total number of steps taken each day
```{r echo = TRUE}

# hist(byday$steps, breaks=61
#      , main = "Total Number of Steps per Day"
#      , col="yellow"
#      , xlab="Day"
#      , ylab = "Steps"
#      , axes = TRUE
#      , plot = TRUE)

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
ggplot(byday, aes(x = day, y = steps)) + geom_bar(stat = "identity")


```

*(A2)*  Calculate and report the mean and median total number of steps taken per day
The mean number of steps per day is `r format(davg, digits = 2, big.mark = ",", decimal.mark = ".")`
and the median number of steps per day is `r format(dmed, digits = 2, big.mark = ",", decimal.mark = ".")`
 
B. What is the average daily activity pattern?
========================================================
*(B1)*  Make a time series plot (i.e.  type = "l" ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

*(B2)*  Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo = TRUE}
plot(davg, ylab="yyyy", xlab = "xxxxx", main = "mainnnn")

ggplot(byinterval, aes(x = interval, y = steps)) + geom_line()

```

 
C. Imputing missing values
========================================================

Note that there are a number of days/intervals where there are missing values (coded as  NA ). The presence of missing days may introduce bias into some calculations or summaries of the data.

*(C1)*  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with  NA s).
Missing values total to `r nMissingCount`
```{r echo = TRUE}


```

*(C2)*  Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Stategy: Replace NA values with the mean, rounded to the nearest whole number, for the the matching interval These values are place in a newly created stepsmod variable combined with the original non NA values.  


*(C3)*  Create a new dataset that is equal to the original dataset but with the missing data filled in.

New Datasetname: activitynew
Here are structures of the orignal and copied objects.

```{r echo = TRUE}
str(activityorig)  #visualize that original object
str(activitynew)  #is the same as new object witouth the NAs
```

*(C4)*  Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r echo = TRUE}

# hist(byday$steps, breaks=61
#      , main = "Total Number of Steps per Day"
#      , col="green"
#      , xlab="Day"
#      , ylab = "Steps"
#      , axes = TRUE
#      , plot = TRUE)

ggplot(byday, aes(x = day, y = steps)) + geom_bar(stat = "identity")

# hist(activitynew$steps 
#      , main="Total Number of Steps per Day with Missing Values Imputed"
#      , xlab ="Total Steps per Day")

#byday <- tapply(activity$steps,activity$date,sum )
# hist(bydaynew, breaks=61
#      , main = "Total Number of Steps per Day (missing values imputed)"
#      , col="green"
#      , xlab="Day"
#      , ylab = "Average Steps"
#      , axes = TRUE
#      , plot = TRUE)

```

 
D. Are there differences in activity patterns between weekdays and weekends?
========================================================

For this part the  weekdays()  function may be of some help here. Use the dataset with the filled-in missing values for this part.
*(D1)*  Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo = TRUE}





```

*(D2)* Panel plot containing a time series plot (i.e.  type = "l" ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```{r echo = TRUE}

#ggplot(byinterval, aes(x = interval, y = steps)) + geom_line() + 
#    facet_wrap(~daytype, nrow = 2)

```

***END****