# Reproducible Research: Peer Assessment 1

# Personal Activity Monitorring Report 

This Knitted report is for the Reproducible Research class Homework Assignment #1. This report is an analysis of Peronal Activity
of the number of steps collected over five minute intervals from a single device/person over a two month period from October through
November of 2012.

## A) Loading and preprocessing the data
Please note that most all of the data processing used to genterate the figures is done in this section. This leaves just the generation of the figures for later in the code.
```{r echo = TRUE}

#opts_chunk$set(echo = TRUE, results = "show", cache = TRUE)

#Download the Activity data if it doesn't already exist
#setwd("C://Users//akeller.HARDEN//Documents//GitHub//RepData_PeerAssessment2")
internetSource <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
fullzipfilename <- file.path(getwd(), "activity.zip")
fullfilename <- file.path(getwd(), "activity.csv")

## note: the following code has been commented out as is not working when Knitting, however, it can be used (uncommented) to run direct from the console in order to download the zip file from the internet
# if (!file.exists(fullzipfilename)){
#   #print("file not found locally, so dowload from website")
#   download.file(internetSource 
#                 , destfile=fullzipfilename)
# }else{
#   #print("data previously downloaded and exists locally")
# }

#unzip contents to get single .csv file, 
# but only if unzipped .csv file doesn't already exist
# assumes activity.zip is in Working Directory
if (!file.exists(fullfilename)){
  unzip(fullzipfilename)
}

#load data into R
if(!exists("activity")){
  activity <- read.csv(fullfilename, header = TRUE, colClasses = c("numeric", "Date", "numeric"))
  #dataset <- read.csv(datasetPath, colClasses = c("numeric", "Date", "numeric"))
} else {print("activity already loaded")}

activityorig <- activity  #copy off original to later compare to newly copied object

#cut

# add a datimetime variable as a POSIXt class which combines date and hour/minutes from interval
activity$datetime <- strptime(
  paste(
    activity$date
    ,activity$intervalmod) 
    ,format="%Y-%m-%d %H%M")

#sum values by Day and by Interval
byday <- tapply(activity$steps,activity$date,sum, na.rm = TRUE )
byday <- aggregate(activity$steps,list(activity$date), sum, na.rm = TRUE )
colnames(byday)[1]<-"day"
colnames(byday)[2]<-"steps"
# summary(byday)

davg <- mean(byday$steps, na.rm=TRUE)
dmed <- median(byday$steps, na.rm=TRUE)

# davg  => 9354.23
# dmed  => 10395

byinterval <- tapply(activityorig$steps,activityorig$interval,sum, na.rm = TRUE )
byinterval <- aggregate(activityorig$steps,list(activityorig$interval), sum,na.rm = TRUE )
colnames(byinterval)[1]<-"interval"
colnames(byinterval)[2]<-"steps"


#byinterval <- tapply(activity$steps,activity$intervalmod,sum, na.rm = TRUE )
#byinterval <- aggregate(activity$steps,list(activity$intervalmod), sum,na.rm = TRUE )
#merge byinterval with stepsmod to replace NULL values in stepsmod
#replaced with average steps for that timeinterval period rounded to the nearest whole numbers
#get interval means and put in a dataframe

#need to count number of missings. Used in output later
nMissingCount <- sum(is.na(activity$steps)
    |is.na(activity$date)
    |is.na(activity$interval)
    )

#nMissingCount  => 2304

#ak question. do I need leading zeros here or not?
#make intervalmod a factor variable
factor(activity$intervalmod)
#pad new intervalmod variable with leading zeros so that it is consistantly HHMM (military hrs)
activity$intervalmod <- sprintf("%04d",activity$interval)

byinterval$intervalmod <- byinterval$interval
factor(byinterval$intervalmod)
byinterval$intervalmod <- sprintf("%04d",byinterval$intervalmod)

#put interval mean in stepsmod where values are NA in steps. otherwise use actual steps 
activity$stepsmod <-  ifelse(
  is.na(activity$steps)
  , round(byinterval$steps[match(activity$intervalmod, byinterval$intervalmod)], digits=0)
  , activity$steps
  )

activitynew <- activity[,c(6,2,3)]
colnames(activitynew)[1] <- "steps" #rename stepsmod back to steps

#ak question: are the empty days overpopulating and should they be included?
bydaynew <- tapply(activitynew$steps,activitynew$date,sum, na.rm = TRUE )
bydaynew <- aggregate(activitynew$steps,list(activitynew$date), sum,na.rm = TRUE )
colnames(bydaynew)[1]<-"day"
colnames(bydaynew)[2]<-"steps"

#byday <- tapply(activity$steps, activity$date, sum)
davgnew <- mean(bydaynew$steps, na.rm=TRUE)
dmednew <- median(bydaynew$steps, na.rm=TRUE)


# davgnew  =>  84188.07
# dmednew  =>  11458

#AverageDailySteps <- davg$steps
# activity$daytype <- ""
# factor(activity$daytype)
activity$daytype <- factor(ifelse (
  weekdays(activity[,"datetime"])=="Saturday"   #if Saturday or Sunday
  | weekdays(activity[,"datetime"])=="Sunday"
  , "Weekend"               #then Weekend
  ,"Weekday"))


```


## B) What is mean total number of steps taken per day?

For now, missing values are ignored in the dataset.
*(B1)*  Here is a histogram of the total number of steps taken each day
```{r echo = TRUE}

# hist(byday$steps, breaks=61
#      , main = "Total Number of Steps per Day"
#      , col="yellow"
#      , xlab="Day"
#      , ylab = "Steps"
#      , axes = TRUE
#      , plot = TRUE)

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
ggplot(byday, aes(x = day, y = steps)) + geom_bar(stat = "identity")


```

*(B2)*  The mean number of steps per day is `r format(davg, digits = 2, big.mark = ",", decimal.mark = ".")`
and the median number of steps per day is `r format(dmed, digits = 2, big.mark = ",", decimal.mark = ".")`

For later...................................

## C) What is the average daily activity pattern?

*(C1)*  Time series plot (i.e.  type = "l" ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo = TRUE}

library(ggplot2)
qplot(byinterval$interval, byinterval$steps, type = "l") + geom_line() + 
    labs(x = "5-min Intervals", y = "Average steps/interval") + theme_bw()

# plot(davg, ylab="yyyy", xlab = "xxxxx", main = "mainnnn")
# 
# ggplot(byinterval, aes(x = interval, y = steps)) + geom_line()

```

```{r echo = TRUE}
max_interval <- byinterval[byinterval$steps == max(byinterval$steps), byinterval$interval] 
# max(byinterval$steps)  => 10927

```

*(C2)*  The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps? is: `r max_interval`.

 
## D. Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as  NA ). The presence of missing days introduces bias into some calculations or summaries of the data.

*(D1)*  The total number of missing values in the dataset (i.e. the total number of rows including NA s) is `r nMissingCount`

*(D2)*  

NA Replacement Stategy: I have replaced NA values with the mean of the matching tiem interval rounded to the nearest whole number. These values are placed in a newly created "stepsmod" variable combined with the originally provided  non-NA values.  

*(D3)*  Create a new dataset that is equal to the original dataset but with the missing data filled in.

The new dataset name is called activitynew.

Here are structures of the orignal and copied objects.

```{r echo = TRUE}
#str(activity)  #visualize that original object
str(activityorig)  #visualize that original object
str(activitynew)  #is the same as new object witout the NAs
```

*(D4)*  Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

The new mean number of steps per day is `r format(davgnew, digits = 2, big.mark = ",", decimal.mark = ".")`
and the new median number of steps per day is `r format(dmednew, digits = 2, big.mark = ",", decimal.mark = ".")`


```{r echo = TRUE}

# hist(byday$steps, breaks=61
#      , main = "Total Number of Steps per Day"
#      , col="green"
#      , xlab="Day"
#      , ylab = "Steps"
#      , axes = TRUE
#      , plot = TRUE)

#ggplot(byday, aes(x = day, y = steps)) + geom_bar(stat = "identity")
ggplot(bydaynew, aes(x = day, y = steps)) + geom_bar(stat = "identity")

# hist(activitynew$steps 
#      , main="Total Number of Steps per Day with Missing Values Imputed"
#      , xlab ="Total Steps per Day")

#byday <- tapply(activity$steps,activity$date,sum )
# hist(bydaynew, breaks=61
#      , main = "Total Number of Steps per Day (missing values imputed)"
#      , col="green"
#      , xlab="Day"
#      , ylab = "Average Steps"
#      , axes = TRUE
#      , plot = TRUE)

```

## E) Are there differences in activity patterns between weekdays and weekends?

For this part the  weekdays()  function may be of some help here. Use the dataset with the filled-in missing values for this part.
*(E1)*  Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo = TRUE}





```
*(E2)* Panel plot containing a time series plot (i.e.  type = "l" ) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```{r echo = TRUE}

#ggplot(byinterval, aes(x = interval, y = steps)) + geom_line() + 
#    facet_wrap(~daytype, nrow = 2)

#start 333 gold maybe
library(plyr)
weekdaysList <- weekdays(activity$date)
weekdaysList[!(weekdaysList %in% c("Saturday", "Sunday"))] <- "weekday"
weekdaysList[(weekdaysList %in% c("Saturday", "Sunday"))] <- "weekend"
#datasetImputed$weekday <- as.factor(weekdaysList)

# dailyStepsByWeekdayImputed <- ddply(datasetImputed, c("interval", "weekday"), 
#     summarize, steps = mean(steps))
# library(lattice)
# xyplot(steps ~ interval | weekday, data = dailyStepsByWeekdayImputed, layout = c(1, 
#     2), type = "l", xlab = "Interval", ylab = "Number of steps")

#almost there. consider plotting aggregate
library(lattice)
xyplot(steps ~ interval | daytype, data = activity, layout = c(1, 
    2), type = "l", xlab = "Interval", ylab = "Number of steps")

```

***END****